### 2017-06-09 Paper/paper.tex
style fixes
- margins
- times font
- parskip, parindent

\documentclass[a4paper,14pt]{extarticle}
\usepackage[left=3cm, right=1.5cm, top=2cm, bottom=2cm]{geometry}
\geometry{a4paper}                   % ... or a4paper or a5paper or ...
\usepackage[T2A]{fontenc}
\usepackage[utf8x]{inputenc}
\usepackage[russian]{babel}
\usepackage{tempora}% for text
\usepackage{newtxmath}% for math
\usepackage{float}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage[colorinlistoftodos]{todonotes}
\renewcommand{\baselinestretch}{1.5}
\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
\usepackage{subfiles}
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}

\begin{document}

\begin{titlepage}
\subfile{sections/0-titlepage}
\end{titlepage}

\tableofcontents
\newpage

\section{Введение}
\subfile{sections/1-introduction}
\newpage

\section{Описание аппаратной части разработанной системы}
\subfile{sections/2-hardware}
\newpage

\section{Описание программного обеспечения}
\subfile{sections/3-software}
\newpage

\section{Заключение}
\subfile{sections/4-conclusion}
\newpage

\subfile{sections/5-bibliography}

\end{document}