### 2019-06-09 docs/master-thesis/paper/sections/9.1-appendix2.tex

% !TEX root = ../main.tex
\documentclass[../main.tex]{subfiles}
\begin{document}
\section{Приложение 2}
\label{appendix2}
Класс AppGUI для модели распространения звука в легких

\begin{lstlisting}
class AppGUI(QtGui.QWidget):
    steps_state = QtCore.pyqtSignal([int])

    def __init__(self):
        super().__init__()

        self.model = LungsModel()

        self.data = self.model.P
        self.observ_slice = np.zeros(self.model.signal_window)
        self.z_slice = self.model.A
        self.y_slice = self.model.B
        self.x_slice = self.model.C

        self.init_ui()
        self.qt_connections()

    def init_ui(self):
        pg.setConfigOption('background', 'w')
        pg.setConfigOption('imageAxisOrder', 'row-major')

        self.z_axis_name = ('Head', 'Feet')
        self.y_axis_name = ('Face', 'Back')
        self.x_axis_name = ('Left Hand', 'Right Hand')

        self.layout = QtGui.QVBoxLayout()

        # self.setGeometry(50, 50, 700, 700)
        self.setWindowTitle('Lungs Model')
        self.l_label = QtGui.QLabel('dt')
        self.h_label = QtGui.QLabel('h')
        self.f_label = QtGui.QLabel('f')

        self.l_spin = pg.SpinBox(value=self.model.l, step=0.01, siPrefix=False, suffix='s')
        self.h_spin = pg.SpinBox(value=self.model.h, step=0.01, siPrefix=False)
        self.f_spin = pg.SpinBox(value=self.model.f, step=1, siPrefix=False)
        self.l_spin.setMaximumWidth(150)
        self.h_spin.setMaximumWidth(150)
        self.f_spin.setMaximumWidth(150)

        self.reset_params_button = QtGui.QPushButton('Defaults')
        self.reinit_button = QtGui.QPushButton('Restart Model')

        self.model_params_layout = QtGui.QHBoxLayout()
        self.model_params_layout.addWidget(self.l_label)
        self.model_params_layout.addWidget(self.l_spin)
        self.model_params_layout.addWidget(self.h_label)
        self.model_params_layout.addWidget(self.h_spin)
        self.model_params_layout.addWidget(self.f_label)
        self.model_params_layout.addWidget(self.f_spin)

        # radio buttons ------------------------------------------------------------
        self.arrays_to_vis = [QtGui.QRadioButton('P'), QtGui.QRadioButton('r'), QtGui.QRadioButton('ro'), QtGui.QRadioButton('c'), QtGui.QRadioButton('K')]
        self.arrays_to_vis[0].setChecked(True)
        # self.radio_layout = QtGui.QHBoxLayout()

        for rad in self.arrays_to_vis:
            self.model_params_layout.addWidget(rad)
            rad.clicked.connect(self.array_to_vis_changed)

        self.model_params_layout.addWidget(self.reset_params_button)
        self.model_params_layout.addWidget(self.reinit_button)

        self.z_slice_label = QtGui.QLabel(f'Z axis [{self.z_axis_name[0]} - {self.z_axis_name[1]}] Slice: {self.z_slice + 1}/{self.data.shape[0]}')
        self.y_slice_label = QtGui.QLabel(f'Y axis [{self.y_axis_name[0]} - {self.y_axis_name[1]}] Slice: {self.y_slice + 1}/{self.data.shape[1]}')
        self.x_slice_label = QtGui.QLabel(f'X axis [{self.x_axis_name[0]} - {self.x_axis_name[1]}] Slice: {self.x_slice + 1}/{self.data.shape[2]}')

        # slices plots ----------------------------------------------------------------

        self.autolevels = True
        self.levels = (0, 100)
        self.glayout = pg.GraphicsLayoutWidget()
        self.glayout.ci.layout.setContentsMargins(0, 0, 0, 0)
        self.glayout.ci.layout.setSpacing(0)
        self.z_slice_img = pg.ImageItem(self.data[self.z_slice, :, :], autoLevels=self.autolevels, levels=self.levels, border=pg.mkPen(color='r', width=3))
        self.y_slice_img = pg.ImageItem(self.data[:, self.y_slice, :], autoLevels=self.autolevels, levels=self.levels, border=pg.mkPen(color='g', width=3))
        self.x_slice_img = pg.ImageItem(self.data[:, :, self.x_slice], autoLevels=self.autolevels, levels=self.levels, border=pg.mkPen(color='b', width=3))
        self.z_slice_plot = self.glayout.addPlot()
        self.y_slice_plot = self.glayout.addPlot()
        self.x_slice_plot = self.glayout.addPlot()
        # self.z_slice_plot.setTitle(f'Z axis [{self.z_axis_name[0]} - {self.z_axis_name[1]}]')
        # self.y_slice_plot.setTitle(f'Y axis [{self.y_axis_name[0]} - {self.y_axis_name[1]}]')
        # self.x_slice_plot.setTitle(f'X axis [{self.x_axis_name[0]} - {self.x_axis_name[1]}]')
        self.z_slice_plot.setAspectLocked() 
        self.y_slice_plot.setAspectLocked() 
        self.x_slice_plot.setAspectLocked() 
        self.z_slice_plot.setMouseEnabled(x=False, y=False)
        self.y_slice_plot.setMouseEnabled(x=False, y=False)
        self.x_slice_plot.setMouseEnabled(x=False, y=False)
        self.z_slice_plot_y_helper1 = self.z_slice_plot.plot([0               , self.data.shape[2] ], [self.y_slice    , self.y_slice      ], pen='g')
        self.z_slice_plot_y_helper2 = self.z_slice_plot.plot([0               , self.data.shape[2] ], [self.y_slice + 1, self.y_slice + 1  ], pen='g')
        self.z_slice_plot_x_helper1 = self.z_slice_plot.plot([self.x_slice    , self.x_slice       ], [0               , self.data.shape[1]], pen='b')
        self.z_slice_plot_x_helper2 = self.z_slice_plot.plot([self.x_slice + 1, self.x_slice + 1   ], [0               , self.data.shape[1]], pen='b')
        self.y_slice_plot_z_helper1 = self.y_slice_plot.plot([0               , self.data.shape[2] ], [self.z_slice    , self.z_slice      ], pen='r')
        self.y_slice_plot_z_helper2 = self.y_slice_plot.plot([0               , self.data.shape[2] ], [self.z_slice + 1, self.z_slice + 1  ], pen='r')
        self.y_slice_plot_x_helper1 = self.y_slice_plot.plot([self.x_slice    , self.x_slice       ], [0               , self.data.shape[0]], pen='b')
        self.y_slice_plot_x_helper2 = self.y_slice_plot.plot([self.x_slice + 1, self.x_slice + 1   ], [0               , self.data.shape[0]], pen='b')
        self.x_slice_plot_z_helper1 = self.x_slice_plot.plot([0               , self.data.shape[1] ], [self.z_slice    , self.z_slice      ], pen='r')
        self.x_slice_plot_z_helper2 = self.x_slice_plot.plot([0               , self.data.shape[1] ], [self.z_slice + 1, self.z_slice + 1  ], pen='r')
        self.x_slice_plot_y_helper1 = self.x_slice_plot.plot([self.y_slice    , self.y_slice       ], [0               , self.data.shape[0]], pen='g')
        self.x_slice_plot_y_helper2 = self.x_slice_plot.plot([self.y_slice + 1, self.y_slice + 1   ], [0               , self.data.shape[0]], pen='g')
        self.z_slice_plot.invertY(True)
        self.y_slice_plot.invertY(True)
        self.x_slice_plot.invertY(True)
        self.z_slice_plot.setLabel('bottom', f'X axis [{self.x_axis_name[0]} - {self.x_axis_name[1]}]')
        self.z_slice_plot.setLabel('left'  , f'Y axis [{self.y_axis_name[1]} - {self.y_axis_name[0]}]')
        self.y_slice_plot.setLabel('bottom', f'X axis [{self.x_axis_name[0]} - {self.x_axis_name[1]}]')
        self.y_slice_plot.setLabel('left'  , f'Z axis [{self.z_axis_name[1]} - {self.z_axis_name[0]}]')
        self.x_slice_plot.setLabel('bottom', f'Y axis [{self.y_axis_name[0]} - {self.y_axis_name[1]}]')
        self.x_slice_plot.setLabel('left'  , f'Z axis [{self.z_axis_name[1]} - {self.z_axis_name[0]}]')
        self.z_slice_plot.addItem(self.z_slice_img)
        self.y_slice_plot.addItem(self.y_slice_img)
        self.x_slice_plot.addItem(self.x_slice_img)
        self.z_slice_img.setRect(pg.QtCore.QRectF(0, 0, self.data.shape[2], self.data.shape[1]))
        self.y_slice_img.setRect(pg.QtCore.QRectF(0, 0, self.data.shape[2], self.data.shape[0]))
        self.x_slice_img.setRect(pg.QtCore.QRectF(0, 0, self.data.shape[1], self.data.shape[0]))
        self.z_slice_img.setZValue(-1)
        self.y_slice_img.setZValue(-1)
        self.x_slice_img.setZValue(-1)

        #--------------------------- signal plots ------------------------
        plots_font = QtGui.QFont()
        fontsize = 9
        plots_font.setPixelSize(fontsize)
        plots_height = 250

        self.source_plot = pg.PlotWidget(title=f'Acoustic Pressure at P[{self.model.A}, {self.model.B}, {self.model.C}] (sound source)')
        self.source_plot.showGrid(x=True, y=True, alpha=0.1)
        self.source_plot.setYRange(-1, 1) # w\o np.log(a)
        # self.fft_widget.setYRange(-15, 0) # w/ np.log(a)
        self.source_plot.getAxis('bottom').setStyle(tickTextOffset = fontsize)
        self.source_plot.getAxis('left').setStyle(tickTextOffset = fontsize)
        self.source_plot.getAxis('bottom').tickFont = plots_font
        self.source_plot.getAxis('left').tickFont = plots_font
        self.source_plot.setMaximumHeight(plots_height)
        self.source_curve = self.source_plot.plot(pen='b')

        self.observ_plot = pg.PlotWidget(title=f'Acoustic Pressure at P[{self.model.oA}, {self.model.oB}, {self.model.oC}]')
        self.observ_plot.showGrid(x=True, y=True, alpha=0.1)
        self.observ_plot.getAxis('bottom').setStyle(tickTextOffset = fontsize)
        self.observ_plot.getAxis('left').setStyle(tickTextOffset = fontsize)
        self.observ_plot.getAxis('bottom').tickFont = plots_font
        self.observ_plot.getAxis('left').tickFont = plots_font
        self.observ_plot.setMaximumHeight(plots_height)
        self.observ_curve = self.observ_plot.plot(pen='r')

        self.observ_slice_plot = pg.PlotWidget(title=f'Acoustic Pressure at P[{self.z_slice}, {self.y_slice}, {self.x_slice}]')
        self.observ_slice_plot.showGrid(x=True, y=True, alpha=0.1)
        self.observ_slice_plot.getAxis('bottom').setStyle(tickTextOffset = fontsize)
        self.observ_slice_plot.getAxis('left').setStyle(tickTextOffset = fontsize)
        self.observ_slice_plot.getAxis('bottom').tickFont = plots_font
        self.observ_slice_plot.getAxis('left').tickFont = plots_font
        self.observ_slice_plot.setMaximumHeight(plots_height)
        self.observ_slice_curve = self.observ_slice_plot.plot(pen=0.8)

        self.plots_layout = QtGui.QHBoxLayout()
        # self.plots_layout.addStretch()
        # self.plots_layout.setMinimumWidth(200)
        self.plots_layout.addWidget(self.source_plot)
        self.plots_layout.addWidget(self.observ_plot)
        self.plots_layout.addWidget(self.observ_slice_plot)

        #----------------------------------------------------------------

        self.step_layout = QtGui.QHBoxLayout()
        self.steps_label = QtGui.QLabel('Number of steps: ')
        self.steps_spin = QtGui.QSpinBox()
        self.steps_spin.setRange(1, 10000)
        self.steps_spin.setValue(50)
        self.steps_spin.setMaximumWidth(100)
        # self.steps_spin.setMaximumSize(100, 50)
        # self.steps_spin.setGeometry(QtCore.QRect(10, 10, 50, 21))
        self.step_button = QtGui.QPushButton('Step')
        # self.step_button.setMaximumSize(100, 50)
        self.step_button.setMaximumWidth(100)
        self.steps_progress_bar = QtGui.QProgressBar()
        self.step_layout.addWidget(self.steps_label)
        self.step_layout.addWidget(self.steps_spin)
        self.step_layout.addWidget(self.step_button)
        self.step_layout.addWidget(self.steps_progress_bar)

        self.z_slice_slider = QtGui.QSlider()
        self.z_slice_slider.setStyleSheet('background-color: rgba(255, 0, 0, 0.2)')
        self.z_slice_slider.setOrientation(QtCore.Qt.Horizontal)
        self.z_slice_slider.setRange(0, self.data.shape[0] - 1)
        self.z_slice_slider.setValue(self.z_slice)
        self.z_slice_slider.setTickPosition(QtGui.QSlider.TicksBelow)
        self.z_slice_slider.setTickInterval(1)

        self.y_slice_slider = QtGui.QSlider()
        self.y_slice_slider.setStyleSheet('background-color: rgba(0, 255, 0, 0.2)')
        self.y_slice_slider.setOrientation(QtCore.Qt.Horizontal)
        self.y_slice_slider.setRange(0, self.data.shape[1] - 1)
        self.y_slice_slider.setValue(self.y_slice)
        self.y_slice_slider.setTickPosition(QtGui.QSlider.TicksBelow)
        self.y_slice_slider.setTickInterval(1)

        self.x_slice_slider = QtGui.QSlider()
        self.x_slice_slider.setStyleSheet('background-color: rgba(0, 0, 255, 0.2)')
        self.x_slice_slider.setOrientation(QtCore.Qt.Horizontal)
        self.x_slice_slider.setRange(0, self.data.shape[2] - 1)
        self.x_slice_slider.setValue(self.x_slice)
        self.x_slice_slider.setTickPosition(QtGui.QSlider.TicksBelow)
        self.x_slice_slider.setTickInterval(1)

        self.layout.addLayout(self.model_params_layout)
        # self.layout.addLayout(self.radio_layout)
        self.layout.addWidget(self.z_slice_label)
        self.layout.addWidget(self.z_slice_slider)
        self.layout.addWidget(self.y_slice_label)
        self.layout.addWidget(self.y_slice_slider)
        self.layout.addWidget(self.x_slice_label)
        self.layout.addWidget(self.x_slice_slider)
        self.layout.addWidget(self.glayout)
        self.layout.addLayout(self.plots_layout)
        # self.layout.addWidget(self.source_plot)
        # self.layout.addWidget(self.observ_plot)
        self.layout.addLayout(self.step_layout)

        self.setLayout(self.layout)

        self.setGeometry(0, 0, 1440, 900)
        # self.setGeometry(0, 0, 1200, 900)
        self.show()

    def qt_connections(self):
        self.step_button.clicked.connect(self.do_steps)
        self.l_spin.valueChanged.connect(self.l_spin_value_changed)
        self.h_spin.valueChanged.connect(self.h_spin_value_changed)
        self.f_spin.valueChanged.connect(self.f_spin_value_changed)
        self.reset_params_button.clicked.connect(self.reset_params)
        self.reinit_button.clicked.connect(self.reinit_model)
        self.z_slice_slider.valueChanged.connect(self.z_slice_slider_changed)
        self.y_slice_slider.valueChanged.connect(self.y_slice_slider_changed)
        self.x_slice_slider.valueChanged.connect(self.x_slice_slider_changed)
        self.steps_state.connect(self.update_steps_progress_bar)

    # def mouseMoved(self, event):
        # print('mouseMoved event')

    def mouseMoveEvent(self, ev):
        print('pp')

    def l_spin_value_changed(self):
        self.model.l = self.l_spin.value()

    def h_spin_value_changed(self):
        self.model.h = self.h_spin.value()

    def f_spin_value_changed(self):
        self.model.f = self.f_spin.value()

    @QtCore.pyqtSlot(int)
    def update_steps_progress_bar(self, current_step):
        self.steps_progress_bar.setValue(current_step / self.steps_spin.value() * 100)
        QApplication.processEvents()

    def reset_params(self):
        self.l_spin.setValue(LungsModel.l_default)
        self.h_spin.setValue(LungsModel.h_default)
        self.f_spin.setValue(LungsModel.f_default)

        self.z_slice = self.model.A
        self.y_slice = self.model.B
        self.x_slice = self.model.C
        self.z_slice_slider.setValue(self.z_slice)
        self.y_slice_slider.setValue(self.y_slice)
        self.x_slice_slider.setValue(self.x_slice)

        self.arrays_to_vis[0].setChecked(True)
        self.array_to_vis_changed()

    def reinit_model(self):
        self.model = LungsModel(self.l_spin.value(), self.h_spin.value(), self.f_spin.value())
        self.array_to_vis_changed()

    def array_to_vis_changed(self):

        mapping = {
            'P' : self.model.P,
            'r' : self.model.r,
            'ro': self.model.ro,
            'c' : self.model.c,
            'K' : self.model.K,
        }

        for r in self.arrays_to_vis:
            if r.isChecked():
                self.data = mapping[r.text()]
                self.z_slice_img.setImage(self.data[self.z_slice      ])
                self.y_slice_img.setImage(self.data[:, self.y_slice, :])
                self.x_slice_img.setImage(self.data[:, :, self.x_slice])

    def print_mean(self):
        # pass
        print(f'slices mean Z, Y, X   {np.mean(self.data[self.z_slice]):8.3e}    {np.mean(self.data[:, self.y_slice, :]):8.3e}    {np.mean(self.data[:, :, self.x_slice]):8.3e}    cube mean: {np.mean(self.data):8.3e}')

    def do_steps(self):
        for i in range(self.steps_spin.value()):
            self.model.step()

            self.z_slice_img.setImage(self.data[self.z_slice      ])
            self.y_slice_img.setImage(self.data[:, self.y_slice, :])
            self.x_slice_img.setImage(self.data[:, :, self.x_slice])

            self.source_curve.setData(self.model.source_signal)
            self.observ_curve.setData(self.model.observ_signal)

            self.observ_slice = np.roll(self.observ_slice, -1)
            self.observ_slice[-1] = self.model.P[self.z_slice, self.y_slice, self.x_slice]
            self.observ_slice_curve.setData(self.observ_slice)

            self.steps_state.emit(i + 1)
            self.print_mean()
        self.steps_state.emit(0)

    def wheelEvent(self, event):
        if self.z_slice_img.sceneBoundingRect().contains(self.glayout.mapFromParent(event.pos())):
            self.z_slice = np.clip(self.z_slice + np.sign(event.angleDelta().y()), 0, self.data.shape[0] - 1) # change bounds 0..N-1 => 1..N
            self.z_slice_slider.setValue(self.z_slice)
        elif self.y_slice_img.sceneBoundingRect().contains(self.glayout.mapFromParent(event.pos())):
            self.y_slice = np.clip(self.y_slice + np.sign(event.angleDelta().y()), 0, self.data.shape[1] - 1) # change bounds 0..N-1 => 1..N
            self.y_slice_slider.setValue(self.y_slice)
        elif self.x_slice_img.sceneBoundingRect().contains(self.glayout.mapFromParent(event.pos())):
            self.x_slice = np.clip(self.x_slice + np.sign(event.angleDelta().y()), 0, self.data.shape[2] - 1) # change bounds 0..N-1 => 1..N
            self.x_slice_slider.setValue(self.x_slice)

    def keyPressEvent(self, event):
        if type(event) == QtGui.QKeyEvent and event.key() == QtCore.Qt.Key_Up:
            self.do_steps()
            #here accept the event and do something
            # self.record_values_button_clicked()
            event.accept()
        else:
            event.ignore()

    def update_observ_slice_plot(self):
        self.observ_slice_plot.getPlotItem().setTitle(f'Acoustic Pressure at P[{self.z_slice}, {self.y_slice}, {self.x_slice}]')
        self.observ_slice = np.zeros(self.model.signal_window)
        self.observ_slice_curve.setData(self.observ_slice)

    def update_slice_helpers_lines(self):
        # self.z_slice_plot_y_helper.setData([0           , self.data.shape[2] ], [self.y_slice, self.y_slice      ])
        # self.z_slice_plot_x_helper.setData([self.x_slice, self.x_slice       ], [0           , self.data.shape[1]])
        # self.y_slice_plot_z_helper.setData([0           , self.data.shape[2] ], [self.z_slice, self.z_slice      ])
        # self.y_slice_plot_x_helper.setData([self.x_slice, self.x_slice       ], [0           , self.data.shape[0]])
        # self.x_slice_plot_z_helper.setData([0           , self.data.shape[1] ], [self.z_slice, self.z_slice      ])
        # self.x_slice_plot_y_helper.setData([self.y_slice, self.y_slice       ], [0           , self.data.shape[0]])
        self.z_slice_plot_y_helper1.setData([0               , self.data.shape[2] ], [self.y_slice    , self.y_slice      ])
        self.z_slice_plot_y_helper2.setData([0               , self.data.shape[2] ], [self.y_slice + 1, self.y_slice + 1  ])
        self.z_slice_plot_x_helper1.setData([self.x_slice    , self.x_slice       ], [0               , self.data.shape[1]])
        self.z_slice_plot_x_helper2.setData([self.x_slice + 1, self.x_slice + 1   ], [0               , self.data.shape[1]])
        self.y_slice_plot_z_helper1.setData([0               , self.data.shape[2] ], [self.z_slice    , self.z_slice      ])
        self.y_slice_plot_z_helper2.setData([0               , self.data.shape[2] ], [self.z_slice + 1, self.z_slice + 1  ])
        self.y_slice_plot_x_helper1.setData([self.x_slice    , self.x_slice       ], [0               , self.data.shape[0]])
        self.y_slice_plot_x_helper2.setData([self.x_slice + 1, self.x_slice + 1   ], [0               , self.data.shape[0]])
        self.x_slice_plot_z_helper1.setData([0               , self.data.shape[1] ], [self.z_slice    , self.z_slice      ])
        self.x_slice_plot_z_helper2.setData([0               , self.data.shape[1] ], [self.z_slice + 1, self.z_slice + 1  ])
        self.x_slice_plot_y_helper1.setData([self.y_slice    , self.y_slice       ], [0               , self.data.shape[0]])
        self.x_slice_plot_y_helper2.setData([self.y_slice + 1, self.y_slice + 1   ], [0               , self.data.shape[0]])

    def z_slice_slider_changed(self):
        self.z_slice = self.z_slice_slider.value()
        self.z_slice_label.setText(f'Z axis [{self.z_axis_name[0]} - {self.z_axis_name[1]}] Slice: {self.z_slice + 1}/{self.data.shape[0]}')
        self.z_slice_img.setImage(self.data[self.z_slice])
        self.print_mean()
        self.update_observ_slice_plot()
        self.update_slice_helpers_lines()

    def y_slice_slider_changed(self):
        self.y_slice = self.y_slice_slider.value()
        self.y_slice_label.setText(f'Y axis [{self.y_axis_name[0]} - {self.y_axis_name[1]}] Slice: {self.y_slice + 1}/{self.data.shape[1]}')
        self.y_slice_img.setImage(self.data[:, self.y_slice, :])
        self.print_mean()
        self.update_observ_slice_plot()
        self.update_slice_helpers_lines()

    def x_slice_slider_changed(self):
        self.x_slice = self.x_slice_slider.value()
        self.x_slice_label.setText(f'X axis [{self.x_axis_name[0]} - {self.x_axis_name[1]}] Slice: {self.x_slice + 1}/{self.data.shape[2]}')
        self.x_slice_img.setImage(self.data[:, :, self.x_slice])
        self.print_mean()
        self.update_observ_slice_plot()
        self.update_slice_helpers_lines()

app = QtGui.QApplication(sys.argv)
# print(sys.argv[1])
gui = AppGUI()
signal.signal(signal.SIGINT, signal.SIG_DFL)
sys.exit(app.exec())

\end{lstlisting}

\newpage
\end{document}